Extracting Information from AGN Variability

### 2017-04-04 Manuscript.tex
Revision 1
Revision 1

% Abstract of the paper
\begin{abstract}
AGN exhibit rapid, high amplitude stochastic flux variations across the entire electromagnetic spectrum on timescales ranging from hours to years. The cause of this variability is poorly understood. We present a Green's Function-based method for using variability to (1) measure the time-scales on which flux perturbations evolve and (2) characterize the driving flux perturbations. We model the observed light curve of an AGN as a linear differential equation driven by stochastic impulses. We analyze the light curve of the \Kepler AGN Zw 229-15 and find that the observed variability behavior can be modeled as a damped harmonic oscillator perturbed by a colored noise process. The model powerspectrum turns over on time-scale $385$~d. On shorter time-scales, the log-powerspectrum slope varies between $2$ and $4$, explaining the behavior noted by previous studies. We recover and identify both the $5.6$~d and $67$~d timescales reported by previous work using the Green's Function of the C-ARMA equation rather than by directly fitting the powerspectrum of the light curve. These are the timescales the time-scale on which flux perturbations grow, and on which flux perturbations decay back to the steady-state flux level respectively. We make the software package \href{https://github.com/AstroVPK/kali}{\textsc{k\={a}l\={i}}} used to study light curves using our method available to the community.
\end{abstract}

@@

Recently, \citet{Kelly14} suggested that stochastic variability in a large class of astrophysical sources may be modelled using Continuous-time AutoRegressive Moving Average (C-ARMA) processes. The attractiveness of using C-ARMA processes to model stochastic phenomenon stems from two factors: (1) C-ARMA processes directly specify property 2, i.e., the correlation structure of the light curve by providing a differential equation that governs how the light curve evolves over time. This alternative to providing a second-order statistic is powerful because it makes the light curve evolution more amenable to physical interpretation as is shown in this work. It also provides some idea of how to replace Gaussian distributions with more generalized $\alpha$-stable distributions that may be truer to reality \citep{BrockwellMarquardt05, BrockwellLinder09}; and (2) flexibility of the shape of their PSD. Since the PSD of a C-ARMA process is a rational function, it can be made arbitrarily complex to model a very wide range of phenomena. This flexibility is achieved by allowing C-ARMA processes to have multiple Autoregressive (quantified by $p$) and Moving Average (quantified by $q$) terms under the constraint $q < p$. If $\chi$ is a small flux perturbation from the mean flux level, then a C-ARMA($p$,$q$) process that models the time evolution of $\chi$ has the general form
\begin{multline}\label{eq:CARMAIntro}
\mathrm{d}^{p}\chi + \alpha_{1} \mathrm{d}^{p-1}\chi + \ldots + \alpha_{p-1} \mathrm{d}\chi + \alpha_{p} \chi = \\ \beta_{0} (\mathrm{d}W) + \beta_{1} \mathrm{d}(\mathrm{d}W) + \ldots + \beta_{p-1} \mathrm{d}^{p-1}(\mathrm{d}W)
\end{multline}
where $W$ is a \textit{Wiener process} and is responsible for introducing stochasticity into the behavior of $\chi$. The $\alpha$ and $\beta$ are constants that govern the behavior of the resulting process and have dimensions of various powers of time. In this paper, we present a Green's function-based method for understanding what a C-ARMA process tells us about accretion physics.

@@

Let the instantaneous flux emitted by an AGN accretion disk be $F(t)$. We may write $F(t) = F_{0} + \chi(t)$ where the $\chi(t)$ are perturbations to the mean flux, $F_{0}$. While the physical origins of AGN variability remain unclear, the phenomenon is known to be non-linear in the X-ray \citep{UMV05}, suggesting that a non-linear equation(s) governs the behavior of $F(t)$. It is not necessarily the case that this equation is physically meaningful i.e. it may be purely phenomenological, arising out of the complex interactions of numerous physical phenomenon. Regardless, if the $\chi(t)$ are small (e.g. $\sim 10 \%$ in the optical), we can \textit{linearise} the non-linear equation that governs $F(t)$ and obtain a \textit{linear} equation for the perturbations $\chi(t)$. In appendix~\ref{sec:LHS}, we show how the linearisation of an arbitrary non-linear equation for the flux $F(t)$ yields the left hand side (LHS) of the C-ARMA equation,
\begin{equation}\label{eq:LHS}
\mathrm{d}^{p}\chi + \alpha_{1} \mathrm{d}^{p-1}\chi + \ldots + \alpha_{p-1} \mathrm{d}\chi + \alpha_{p} \chi = u.
\end{equation}
Here, $u$ are input impulses that drive the behavior of the flux perturbations, $\chi$.

The Green's function of equation~\eqref{eq:LHS} can provide insights into how accretion physics works. Conceptually, the Green's function of a linear system characterizes the impulse response of the system. Particular solutions are found by linear superposition (integration) of the Green's function over all the impulses that drive the system \citep{PanditWu}. Here, the Green's function quantifies how a unit flux perturbation evolves after it is triggered. By comparing the Green's function to the expected behaviour of perturbations, we can gain insight into accretion physics. The form of the Green's function for equation~\eqref{eq:LHS} is given by
\begin{equation}\label{eq:GFSol}
G(t) = \sum_{k = 1}^{p} c_{k} \mathrm{e}^{\rho_{k}t},
\end{equation}
where the $\rho_{k}$ are the roots of the autoregressive polynomial of equation~\eqref{eq:ARCharPoly} and the $c_{k}$ are constants that can be determined as stated in appendix~\ref{sec:GFComputation}. Using Euler's formula, it is clear that the real part of each $\rho_{k}$ corresponds to an exponential decay timescale $\tau_{k} = 1/\Re(\rho_{k})$ while the complex part of each $\rho_{k}$ produces modulating oscillations with period $T_{k} = 2\upi/\Im(\rho_{k})$.

The cause of perturbations, i.e. variability, in the flux emanating from an AGN is unknown, but we may construct a very general mathematical model for the form of the flux perturbations based on physical and statistical arguments. The region of space surrounding the central engine is vast. The inner edge of the accretion disk is thought to begin a few $r_{G} \sim 20$~AU ($M_{BH} = 10^{9} M_{\sun}$) away from the event horizon, while the outer edge is thought to lie at a distance of a few parsec from the event horizon. Contributions to the deviations in the flux most likely simultaneously arise from many individual locations in this system. Under the assumption that the distribution of energy released by each individual event is drawn from a distribution with finite second-moment, the central-limit-theorem assures us that the total energy introduced into the system at any instant will be drawn from a normal distribution, i.e., some form of Weiner noise. Due to the non-linearity of the physics involved, it is quite likely that correlations exist between the energy released into the disk at different moments in time. Therefore, we shall model the correlations by treating the driving process $u$ in equation~\eqref{eq:LHS} as a correlated Gaussian noise process. In appendix~\ref{sec:RHS}, we argue why the natural mathematical form of the RHS should be given by
\begin{equation}\label{eq:RHS}
u = \beta_{0} (\mathrm{d}W) + \beta_{1} \mathrm{d}(\mathrm{d}W) + \ldots + \beta_{q-1} \mathrm{d}^{q-1}(\mathrm{d}W) + \beta_{q} \mathrm{d}^{q}(\mathrm{d}W),
\end{equation}
with $q < p$. This form for the input stochastic impulses can be characterized by the power spectral density (PSD) of the impulses given by
\begin{equation}\label{eq:ImpulsePSD}
S_{uu}(\nu) = \frac{1}{2 \upi } |\beta(2 \upi \imath \nu)|^{2},
\end{equation}
where $\beta(z)$ is the moving average polynomial of equation~\eqref{eq:MACharPoly}. The PSD of the stochastic impulses tells about the timescales on which the stochastic impulses occur. Combining equation~\eqref{eq:LHS} with equation~\eqref{eq:RHS} yields the original C-ARMA process equation~\eqref{eq:CARMAIntro}. A C-ARMA process is a linear differential equation perturbed/driven by a random process. \textit{We see that C-ARMA processes attempt to describe AGN luminosity variations in terms of a linear system driven by colored noise}.

@@

\section[\textsc{k\={a}l\={i}}]{\textsc{k\={a}l\={i}}: Software for the C-ARMA Analysis of a Light Curve}\label{sec:Steps}

We have implemented a fully parallelized and vectorized software package, \textsc{k\={a}l\={i}} (named jointly after the Hindu goddess of time, change, and power and also as an acronym for KArma LIbrary), to analyze stochastic light curves using the C-ARMA process presented in equation~\eqref{eq:CARMAIntro}. \textsc{k\={a}l\={i}} is implemented in the \textsc{c++} programming language with \textsc{Python} language bindings for ease of use. \textsc{k\={a}l\={i}} may be obtained from \url{https://github.com/AstroVPK/libcarma} and is installable on Linux variants and Mac OSX using the install scripts provided. Appendix~\ref{sec:kali} describes the mathematics implemented in \textsc{k\={a}l\={i}} that are used to perform the Bayesian Markov Chain Monte-Carlo (MCMC) inferencing of a stochastic light curve.

Before performing a C-ARMA analysis, it should first be determined if a C-ARMA model is appropriate. A C-ARMA process may be unsuitable if the light curve exhibits non-stationary behaviour, i.e., (1) the appearance of the light curve changes dramatically at some point in the light curve or (2) the light curve has a marked trend. A C-ARMA analysis may also be unsuitable if the light curve has flares (as in the case of the blazar of \citealp{Edelson13}). C-ARMA models are also unsuitable if the variations in the flux are very large compared to the flux. A C-ARMA model is suitable if the sample autocorrelation and partial autocorrelation functions of the light curve exponentially decay to below the significance level for a pure white noise process \citep{BrockwellDavisITSF}.

@@

We use our package, \textsc{k\={a}l\={i}} to fit the observed light curve of Zw 229-15 to C-ARMA($p$,$q$) processes with $1 \leq p \leq 10$ and $0 \leq q < p$. In addition to using our package \textsc{k\={a}l\={i}}, we performed the same analysis using the \textsc{c++} and \textsc{Python} library \textsc{carma\_pack} of \citet{Kelly14}. While both codes produced similar numerical results, \textsc{k\={a}l\={i}} runs about $50 \times$ faster due to our extensive usage of the high performance Intel Math Kernel Library (MKL). Prior to the analysis, the light curve of Zw 229-15 was corrected to the rest-frame of the galaxy at $z = 0.0275$ by scaling $\delta t$ by $1/(1+z)$. No binning of the light curve was performed. Instead of adopting the \Kepler error estimates, \textsc{carma\_pack} assumes that the variance of the observation noise may be inaccurate and sets $\sigma^{2}_{N,\mathrm{true}} = M_{\mathrm{Err}} \sigma^{2}_{N,\mathrm{stated}}$ for each observation. The multiplicative scaling factor $M_{\mathrm{Err}}$ is treated as a fit parameter, i.e., all purely white noise in the data is assumed to originate in process of observation rather than be intrinsic to the signal. We find that the median value of the multiplicative scaling factor is $M_{\mathrm{Err}} = 1.30$. The observation noise levels inferred from the pixel noise statistics of the \Kepler CCDs appear to underestimate the true observation noise level in the signal by $\sim 14$ percent at the $95$-percent confidence level.

@@

\begin{figure}
	\includegraphics[width=\columnwidth]{images/Zw229-15_mockLC.jpg}
    \caption{Simulated light curves with parameters drawn from the MCMC estimates of the C-ARMA($2$,$1$) parameters of Zw 229-15. The true light curve of Zw 229-15 is also plotted. Can the true light curve be distinguished by eye?}
    \label{fig:Zw229-15_mock}
\end{figure}

If the C-ARMA model is not estimated correctly, mock light curves generated using the model will not resemble the observed light curve. To visually test the C-ARMA($2$,$1$) model inferred by us, we randomly selected sets of model parameter values, i.e., sets of values of $\alpha_{1}$, $\alpha_{2}$, $\beta_{1}$, and $\beta_{2}$, and generated simulated light curves using these parameters with the same sampling and noise levels as the true light curve. Figure~\ref{fig:Zw229-15_mock} shows a sample of three such mock light curves along with the true light curve (orange). Visually all four light curves look very similar suggesting that the C-ARMA($2$,$1$) is capturing most of the information in the light curve.

The light curve of Zw 229-15 observed by \Kepler is well described by a C-ARMA($2$,$1$) process suggesting that a $2^{\mathrm{nd}}$-order (probably non-linear) differential equation may be responsible for smoothing out flux perturbations over time. The \Kepler data suggest that the impulses to the flux that drive the perturbations have equal power on time-scales ranging from days to years. Perturbations to the flux behave like over-damped harmonic oscillators---the perturbation peaks in intensity at $\sim 5.6$~d after which it gradually decays in intensity with an $\mathrm{e}$-folding time-scale of $\sim 69$~d. On very short time-scales ($\sim 1$~d) there are indications that the impulse powerspectrum begins to rise with log-PSD slope $2$ which may be due to the thermal motion of the accretion disk plasma. However, the observation noise level makes it difficult to probe accretion physics on these time-scales. We discuss the effect of the \Kepler noise characteristics on our results in the next section.

@@

We demonstrate that the Continuous-time AutoRegressive Moving Average (C-ARMA) model of \citet{Kelly14} results from linearization of arbitrary non-linear differential equations for the flux emitted by the AGN accretion disk. Small perturbations of the total flux emitted by an AGN can be modelled in the linear-regime as a linear differential equation driven by noise or a C-ARMA process. Such processes consist of an $n^{\mathrm{th}}$-order linear differential equation for the flux perturbation (LHS) stochastically driven by a linear combination of Wiener increments (RHS). The driving impulses may characterize complex MHD processes such as MRI turbulence. We show how insights into the variability-driving physics can be obtained by examining the PSD of the flux impulses. We propose that the homogenized linear differential equation corresponding to the C-ARMA process governs how flux perturbations evolve after they have been generated. We show how the evolution of the flux perturbations can be characterized by the Green's function of the linear-differential equation.

@@

C-ARMA analysis techniques offer tremendous promise for studying AGN \& BHXRB variability. The C-ARMA formalism can be extended to model different accretion states using formalisms such the Threshold C-ARMA processes of \citet{DimensionEstimationBrockwell}. C-ARMA processes can also be extended to non-simultaneous multi-wavelength observations to probe the connection between the variability observed in different bands. Numerical simulations of accretion disks are already able to produce time-domain synthetic spectra \citep{Schnittman13b} and will soon robustly include radiative transfer in the modelling \citep{FragileAccretion}. C-ARMA models may be useful in comparing the synthetic spectra and light curves produced by numerical models of the accretion disk and winds to observational results. Moving forward, more sophisticated techniques such as the particle filter may be able to directly infer the underlying non-linear stochastic differential equations governing variability without the need for linearization \citep{Hanif15}. Tools for performing C-ARMA analysis are provided in the \textsc{c++} and \textsc{Python} package \textsc{k\={a}l\={i}} and can be obtained from \url{https://github.com/AstroVPK/libcarma}.

@@

We acknowledge support from NASA grant NNX14AL56G. This paper includes data collected by the Kepler mission. Funding for the Kepler mission is provided by the NASA Science Mission directorate. Some of the data presented in this paper were obtained from the Mikulski Archive for Space Telescopes (MAST). STScI is operated by the Association of Universities for Research in Astronomy, Inc., under NASA contract NAS5-26555. Support for MAST for non-HST data is provided by the NASA Office of Space Science via grant NNX09AF08G and by other grants and contracts. \textsc{k\={a}l\={i}} uses the optimization libraries provided by Steven G. Johnson, The NLopt nonlinear-optimization package, http://ab-initio.mit.edu/nlopt. We wish to thank Jack O'Brien, Jackie Moreno, Nate B. Lust and Adam Lidz for their input and discussions.

@@

\section[\textsc{k\={a}l\={i}}]{How Does \textsc{k\={a}l\={i}} Infer C-ARMA Parameters}\label{sec:kali}

@@

We represent the C-ARMA process in continuous-time state-space form \citep{Wiberg,Friedland} as
\begin{equation}\label{eq:StateEqn}
\mathrm{d}\mathbfit{x} = \mathbfss{A}\mathbfit{x}\mathrm{d}t + \mathbfit{B}\mathrm{d}w,
\end{equation}
and
\begin{equation}\label{eq:ObsEqn}
\chi(t) = \mathbfss{H(t)}\mathbfit{x} + v,
\end{equation}
where $\mathbfit{x}$ is the system-state vector, $W$ is the Wiener process of section~\ref{sec:RHS}, $v \sim \mathcal{N}(0,\sigma^{2}_{N})$ is the measurement noise, and $\chi$ is the (noisy) observed flux at time $t$. The matrix $\mathbfss{A}$ is known as the `state-transition matrix', while the matrix $\mathbfss{H}$ is known as the `observation matrix'. There are infinitely many ways of writing a C-ARMA process in state-space form--we shall discuss the two most relevant \citep{Denham74,Kailath74} because of their use in common packages for C-ARMA inferencing such as \textsc{k\={a}l\={i}} (this work) and \textsc{carma\_pack} \citep{Kelly14}. All other forms may be obtained from these `canonical' forms via similarity transforms.

>>>>

% Abstract of the paper
\begin{abstract}
AGN exhibit rapid, high amplitude stochastic flux variations across the entire electromagnetic spectrum on timescales ranging from hours to years. The cause of this variability is poorly understood. We present a Green's Function-based method for using variability to (1) measure the time-scales on which flux perturbations evolve and (2) characterize the driving flux perturbations. We model the observed light curve of an AGN as a linear differential equation driven by stochastic impulses. We analyze the light curve of the \Kepler AGN Zw 229-15 and find that the observed variability behavior can be modeled as a damped harmonic oscillator perturbed by a colored noise process. The model powerspectrum turns over on time-scale $385$~d. On shorter time-scales, the log-powerspectrum slope varies between $2$ and $4$, explaining the behavior noted by previous studies. We recover and identify both the $5.6$~d and $67$~d timescales reported by previous work using the Green's Function of the C-ARMA equation rather than by directly fitting the powerspectrum of the light curve. These are the timescales on which flux perturbations grow, and on which flux perturbations decay back to the steady-state flux level respectively. We make the software package \href{https://github.com/AstroVPK/kali}{\textsc{k\={a}l\={i}}} used to study light curves using our method available to the community.
\end{abstract}

@@

Recently, \citet{Kelly14} suggested that stochastic variability in a large class of astrophysical sources may be modelled using Continuous-time AutoRegressive Moving Average (C-ARMA) processes. The attractiveness of using C-ARMA processes to model stochastic phenomenon stems from two factors: (1) C-ARMA processes directly specify property 2, i.e., the correlation structure of the light curve by providing a differential equation that governs how the light curve evolves over time. This alternative to providing a second-order statistic is powerful because it makes the light curve evolution more amenable to physical interpretation as is shown in this work. It also provides some idea of how to replace Gaussian distributions with more generalized $\alpha$-stable distributions that may be truer to reality \citep{BrockwellMarquardt05, BrockwellLinder09}; and (2) flexibility of the shape of their PSD. Since the PSD of a C-ARMA process is a rational function, it can be made arbitrarily complex to model a very wide range of phenomena. This flexibility is achieved by allowing C-ARMA processes to have multiple Autoregressive (quantified by $p$) and Moving Average (quantified by $q$) terms under the constraint $q < p$. If $\chi$ is a small flux perturbation from the mean flux level, then a C-ARMA($p$,$q$) process that models the time evolution of $\chi$ has the general form
\begin{multline}\label{eq:CARMAIntro}
\mathrm{d}^{p}\chi + \alpha_{1} \mathrm{d}^{p-1}\chi + \ldots + \alpha_{p-1} \mathrm{d}\chi + \alpha_{p} \chi = \\ \beta_{0} (\mathrm{d}W) + \beta_{1} \mathrm{d}(\mathrm{d}W) + \ldots + \beta_{p-1} \mathrm{d}^{p-1}(\mathrm{d}W)
\end{multline}
where $W$ is a \textit{Wiener process} and is responsible for introducing stochasticity into the behavior of $\chi$. The $\alpha$ and $\beta$ are constants that govern the behavior of the resulting process and have dimensions of various powers of time. Several studies have found that C-ARMA processes model AGN variability well \citep{Simm16, Caplar17}. In this paper, we present a Green's function-based method for understanding what a C-ARMA process tells us about accretion physics.

@@

We present motivations for using C-ARMA processes to model AGN variability. While the physical origins of AGN variability remain unclear, the phenomenon is thought to arise in the accretion disk of the AGN. Radiation from the accretion disk is further modulated by matter in the region surrounding the disk, forming a complex system. The flux, $F(t)$, emitted by the system as a whole is known to vary non-linearly, with stronger variability being observed when the object is brighter \citep{UMV05}. Assume that there exists some stochastic differential equation that phenomenologically models $F(t)$ over the duration that the AGN system is observed for. The non-linearity observed in AGN variability suggests that any phenomenological model of variability be non-linear. However if the variations in the flux are small, such as are observed in the optical, variability can be modeled as perturbations; we may write $F(t) = F_{0} + \chi(t)$ where the $\chi(t)$ are perturbations to the steady-state flux $F_{0}$. We can then \textit{linearise} the non-linear equation that governs $F(t)$ and obtain a \textit{linear} equation for the perturbations $\chi(t)$. In appendix~\ref{sec:LHS}, we show how the linearisation of an arbitrary non-linear equation for the flux $F(t)$ yields the left hand side (LHS) of the C-ARMA equation,
\begin{equation}\label{eq:LHS}
\mathrm{d}^{p}\chi + \alpha_{1} \mathrm{d}^{p-1}\chi + \ldots + \alpha_{p-1} \mathrm{d}\chi + \alpha_{p} \chi = u.
\end{equation}
Here, $u$ are input impulses that drive the behavior of the flux perturbations, $\chi$.

The impulse response of the system is characterized by the Green's function of equation~\eqref{eq:LHS}. Particular solutions are found by linear superposition (integration) of the Green's function over all the impulses that drive the system \citep{PanditWu}. Here, the Green's function quantifies how a unit flux perturbation evolves after it is triggered. By comparing the observed Green's function to the expected behaviour of perturbations from theoretical considerations, we can gain insight into accretion physics. The form of the Green's function for equation~\eqref{eq:LHS} is given by
\begin{equation}\label{eq:GFSol}
G(t) = \sum_{k = 1}^{p} c_{k} \mathrm{e}^{\rho_{k}t},
\end{equation}
where the $\rho_{k}$ are the roots of the autoregressive polynomial of equation~\eqref{eq:ARCharPoly} and the $c_{k}$ are constants that can be determined as stated in appendix~\ref{sec:GFComputation}. Using Euler's formula, it is clear that the real part of each $\rho_{k}$ corresponds to an exponential decay timescale $\tau_{k} = 1/\Re(\rho_{k})$ while the complex part of each $\rho_{k}$ produces modulating oscillations with period $T_{k} = 2\upi/\Im(\rho_{k})$.

The cause of perturbations, i.e. variability, in the flux emanating from an AGN is unknown; the C-ARMA model of \citep{Kelly14} chooses a very general mathematical model for the form of the flux perturbations, %based on physical and statistical arguments. The region of space surrounding the central engine is vast. The inner edge of the accretion disk is thought to begin a few $r_{G} \sim 20$~AU ($M_{BH} = 10^{9} M_{\sun}$) away from the event horizon, while the outer edge is thought to lie at a distance of a few parsec from the event horizon. Contributions to the deviations in the flux most likely simultaneously arise from many individual locations in this system. Under the assumption that the distribution of energy released by each individual event is drawn from a distribution with finite second-moment, the central-limit-theorem assures us that the total energy introduced into the system at any instant will be drawn from a normal distribution, i.e., some form of Weiner noise. Due to the non-linearity of the physics involved, it is quite likely that correlations exist between the energy released into the disk at different moments in time. Therefore, we choose to model the correlations by treating the driving process $u$ in equation~\eqref{eq:LHS} as
a correlated Gaussian noise process given by
\begin{equation}\label{eq:RHS}
u = \beta_{0} (\mathrm{d}W) + \beta_{1} \mathrm{d}(\mathrm{d}W) + \ldots + \beta_{q-1} \mathrm{d}^{q-1}(\mathrm{d}W) + \beta_{q} \mathrm{d}^{q}(\mathrm{d}W),
\end{equation}
with $q < p$. In appendix~\ref{sec:RHS}, we present an overview of the mathematics of Weiner processes and increments and discuss why it is necessary to impose the condition that $q < p$. Alternatives exist to the C-ARMA process choice of driving impulses such as models in which the magnitude of the driving impulses is determined by some function of the current value of the flux such as in the Cox-Ingersoll-Ross model \citep{CoxIngersollRoss85}.

The form for the input stochastic impulses can be characterized by the power spectral density (PSD) of the impulses given by
\begin{equation}\label{eq:ImpulsePSD}
S_{uu}(\nu) = \frac{1}{2 \upi } |\beta(2 \upi \imath \nu)|^{2},
\end{equation}
where $\beta(z)$ is the moving average polynomial of equation~\eqref{eq:MACharPoly}. The PSD of the stochastic impulses tells about the timescales on which the stochastic impulses occur. Combining equation~\eqref{eq:LHS} with equation~\eqref{eq:RHS} yields the original C-ARMA process equation~\eqref{eq:CARMAIntro}. A C-ARMA process is a linear differential equation perturbed/driven by a random process. \textit{We see that C-ARMA processes attempt to describe AGN luminosity variations in terms of a linear system driven by colored noise}.

@@

\section[\href{https://github.com/AstroVPK/kali}{\textsc{k\={a}l\={i}}}]{\href{https://github.com/AstroVPK/kali}{\textsc{k\={a}l\={i}}}: Software for the C-ARMA Analysis of a Light Curve}\label{sec:Steps}

We have implemented a fully parallelized and vectorized software package, \href{https://github.com/AstroVPK/kali}{\textsc{k\={a}l\={i}}} (named jointly after the Hindu goddess of time, change, and power and also as an acronym for KArma LIbrary), to analyze stochastic light curves using the C-ARMA process presented in equation~\eqref{eq:CARMAIntro}. \href{https://github.com/AstroVPK/kali}{\textsc{k\={a}l\={i}}} is implemented in the \textsc{c++} programming language with \textsc{Python} language bindings for ease of use. \href{https://github.com/AstroVPK/kali}{\textsc{k\={a}l\={i}}} may be obtained from \url{https://github.com/AstroVPK/libcarma} and is installable on Linux variants and Mac OSX using the install scripts provided. Appendix~\ref{sec:kali} describes the mathematics implemented in \href{https://github.com/AstroVPK/kali}{\textsc{k\={a}l\={i}}} that are used to perform the Bayesian Markov Chain Monte-Carlo (MCMC) inferencing of a stochastic light curve.

Before performing a C-ARMA analysis, it should first be determined if a C-ARMA process is an appropriate model for the light curve of the AGN. A C-ARMA process may be unsuitable if the light curve exhibits non-stationary behaviour, i.e., (1) the powerspectrum of the light curve changes over the course of the light curve at the $2$-$\sigma$ level as determined by estimating the powerspectrum of short segments of the light curve; or (2) the light curve has a marked linear trend as determined from a simple regression test. A C-ARMA analysis may also be unsuitable if the light curve exhibits sudden, large amplitude-short duration changes in flux i.e. `flares' (as in the case of the blazar of \citealp{Edelson13}). A C-ARMA process would have to generate a series of positive large amplitude variations followed by a similar series of negative variations to produce `flare'-like behavior which is mathematically highly unlikely to occur. C-ARMA processes are theoretically able to produce any value for the flux variation, including values that would result in \textit{negative} total flux. Therefore, C-ARMA processes are unsuitable models for AGN variability if the flux variations are very large compared to the flux as this may increases the likelihood of producing an unphysical negative total flux to an unacceptable level. A C-ARMA model is suitable if the sample autocorrelation and partial autocorrelation functions of the light curve exponentially decay to below the $2$-$\sigma$ significance level for a pure white noise process \citep{BrockwellDavisITSF}.

@@

We use our package, \href{https://github.com/AstroVPK/kali}{\textsc{k\={a}l\={i}}} to fit the observed light curve of Zw 229-15 to C-ARMA($p$,$q$) processes with $1 \leq p \leq 10$ and $0 \leq q < p$. In addition to using our package \href{https://github.com/AstroVPK/kali}{\textsc{k\={a}l\={i}}}, we performed the same analysis using the \textsc{c++} and \textsc{Python} library \textsc{carma\_pack} of \citet{Kelly14}. While both codes produced similar numerical results, \href{https://github.com/AstroVPK/kali}{\textsc{k\={a}l\={i}}} runs about $50 \times$ faster due to our extensive usage of the high performance Intel Math Kernel Library (MKL). Prior to the analysis, the light curve of Zw 229-15 was corrected to the rest-frame of the galaxy at $z = 0.0275$ by scaling $\delta t$ by $1/(1+z)$. No binning of the light curve was performed. Instead of adopting the \Kepler error estimates, \textsc{carma\_pack} assumes that the variance of the observation noise may be inaccurate and sets $\sigma^{2}_{N,\mathrm{true}} = M_{\mathrm{Err}} \sigma^{2}_{N,\mathrm{stated}}$ for each observation. The multiplicative scaling factor $M_{\mathrm{Err}}$ is treated as a fit parameter, i.e., all purely white noise in the data is assumed to originate in process of observation rather than be intrinsic to the signal. We find that the median value of the multiplicative scaling factor is $M_{\mathrm{Err}} = 1.30$. The observation noise levels inferred from the pixel noise statistics of the \Kepler CCDs appear to underestimate the true observation noise level in the signal by $\sim 14$ percent at the $95$-percent confidence level.

@@

%\begin{figure}
%	\includegraphics[width=\columnwidth]{images/Zw229-15_mockLC.jpg}
%    \caption{Simulated light curves with parameters drawn from the MCMC estimates of the C-ARMA($2$,$1$) parameters of Zw 229-15. The true light curve of Zw 229-15 is also plotted. Can the true light curve be distinguished by eye?}
%    \label{fig:Zw229-15_mock}
%\end{figure}

%If the C-ARMA model is not estimated correctly, mock light curves generated using the model will not resemble the observed light curve. To visually test the C-ARMA($2$,$1$) model inferred by us, we randomly selected sets of model parameter values, i.e., sets of values of $\alpha_{1}$, $\alpha_{2}$, $\beta_{1}$, and $\beta_{2}$, and generated simulated light curves using these parameters with the same sampling and noise levels as the true light curve. Figure~\ref{fig:Zw229-15_mock} shows a sample of three such mock light curves along with the true light curve (orange). Visually all four light curves look very similar suggesting that the C-ARMA($2$,$1$) is capturing most of the information in the light curve.

The light curve of Zw 229-15 observed by \Kepler is well described by a C-ARMA($2$,$1$) process suggesting that a $2^{\mathrm{nd}}$-order differential equation may be responsible for smoothing out flux perturbations over time. The \Kepler data suggest that the impulses to the flux that drive the perturbations have equal power on time-scales ranging from days to years. Perturbations to the flux behave like over-damped harmonic oscillators---the perturbation peaks in intensity at $\sim 5.6$~d after which it gradually decays in intensity with an $\mathrm{e}$-folding time-scale of $\sim 69$~d. On very short time-scales ($\sim 1$~d) there are indications that the impulse powerspectrum begins to rise with log-PSD slope $2$ which may be due to the thermal motion of the accretion disk plasma. However, the observation noise level makes it difficult to probe accretion physics on these time-scales. We discuss the effect of the \Kepler noise characteristics on our results in the next section.

@@

We demonstrate that the Continuous-time AutoRegressive Moving Average (C-ARMA) model of \citet{Kelly14} results from linearization of arbitrary phenomenological non-linear differential equations for the flux emitted by the AGN accretion disk. Small perturbations of the total flux emitted by an AGN can be modelled in the linear-regime as a linear differential equation driven by noise or a C-ARMA process. Such processes consist of an $n^{\mathrm{th}}$-order linear differential equation for the flux perturbation (LHS) stochastically driven by a linear combination of Wiener increments (RHS). The driving impulses may characterize complex MHD processes such as MRI turbulence. We show how insights into the variability-driving physics can be obtained by examining the PSD of the flux impulses. We propose that the homogenized linear differential equation corresponding to the C-ARMA process governs how flux perturbations evolve after they have been generated. We show how the evolution of the flux perturbations can be characterized by the Green's function of the linear-differential equation.

@@

C-ARMA analysis techniques offer tremendous promise for studying AGN \& BHXRB variability. The C-ARMA formalism can be extended to model different accretion states using formalisms such the Threshold C-ARMA processes of \citet{DimensionEstimationBrockwell}. C-ARMA processes can also be extended to non-simultaneous multi-wavelength observations to probe the connection between the variability observed in different bands. Numerical simulations of accretion disks are already able to produce time-domain synthetic spectra \citep{Schnittman13b} and will soon robustly include radiative transfer in the modelling \citep{FragileAccretion}. C-ARMA models may be useful in comparing the synthetic spectra and light curves produced by numerical models of the accretion disk and winds to observational results. Moving forward, more sophisticated techniques such as the particle filter may be able to directly infer the underlying non-linear stochastic differential equations governing variability without the need for linearization \citep{Hanif15}. Tools for performing C-ARMA analysis are provided in the \textsc{c++} and \textsc{Python} package \href{https://github.com/AstroVPK/kali}{\textsc{k\={a}l\={i}}} and can be obtained from \url{https://github.com/AstroVPK/libcarma}.

@@

We acknowledge support from NASA grant NNX14AL56G. This paper includes data collected by the Kepler mission. Funding for the Kepler mission is provided by the NASA Science Mission directorate. Some of the data presented in this paper were obtained from the Mikulski Archive for Space Telescopes (MAST). STScI is operated by the Association of Universities for Research in Astronomy, Inc., under NASA contract NAS5-26555. Support for MAST for non-HST data is provided by the NASA Office of Space Science via grant NNX09AF08G and by other grants and contracts. \href{https://github.com/AstroVPK/kali}{\textsc{k\={a}l\={i}}} uses the optimization libraries provided by Steven G. Johnson, The NLopt nonlinear-optimization package, http://ab-initio.mit.edu/nlopt. We wish to thank Jack O'Brien, Jackie Moreno, Nate B. Lust and Adam Lidz for their input and discussions.

@@

\section[\href{https://github.com/AstroVPK/kali}{\textsc{k\={a}l\={i}}}]{How Does \href{https://github.com/AstroVPK/kali}{\textsc{k\={a}l\={i}}} Infer C-ARMA Parameters}\label{sec:kali}

@@

We represent the C-ARMA process in continuous-time state-space form \citep{Wiberg,Friedland} as
\begin{equation}\label{eq:StateEqn}
\mathrm{d}\mathbfit{x} = \mathbfss{A}\mathbfit{x}\mathrm{d}t + \mathbfit{B}\mathrm{d}w,
\end{equation}
and
\begin{equation}\label{eq:ObsEqn}
\chi(t) = \mathbfss{H(t)}\mathbfit{x} + v,
\end{equation}
where $\mathbfit{x}$ is the system-state vector, $W$ is the Wiener process of section~\ref{sec:RHS}, $v \sim \mathcal{N}(0,\sigma^{2}_{N})$ is the measurement noise, and $\chi$ is the (noisy) observed flux at time $t$. The matrix $\mathbfss{A}$ is known as the `state-transition matrix', while the matrix $\mathbfss{H}$ is known as the `observation matrix'. There are infinitely many ways of writing a C-ARMA process in state-space form--we shall discuss the two most relevant \citep{Denham74,Kailath74} because of their use in common packages for C-ARMA inferencing such as \href{https://github.com/AstroVPK/kali}{\textsc{k\={a}l\={i}}} (this work) and \textsc{carma\_pack} \citep{Kelly14}. All other forms may be obtained from these `canonical' forms via similarity transforms.
